### 2019-03-26 abstract/abstract.tex
Add in abstract shell with sections

%        File: abstract.tex
%     Created: Tue Mar 26 06:00 AM 2019 E
% Last Change: Tue Mar 26 06:00 AM 2019 E
%
\documentclass{article}
\begin{document}

\title{CS287 Final Project Abstract}
\author{Francisco Rivera \and Jiafeng (Kevin) Chen \and Yufeng Ling}
\date{\today}

\maketitle

\section{Area}

% TODO: write up area

\section{Papers}

% TODO: bring in the two papers and find a third

\section{Time}

We have already signed up for Monday April 22.

\section{Baseline}

% TODO: this is the hardest thing we need to decide

\end{document}